---
title: "At5_MynaMorphology"
author: "K C STUART"
date: "July 2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
setwd("C:/Users/kstu465/OneDrive - The University of Auckland/Desktop/RESEARCH/PROJECTS/At5_MynaMorphology")

knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(ggplot2,dplyr)

```

## Load, filter, and check data

```{r, include = FALSE}
Myna_Pheno <- read.csv("mynas_confirmed_ROM_Kat_Edit.csv",stringsAsFactors=TRUE,sep=",")
str(Myna_Pheno)

#filter
Myna_Pheno_filter <- Myna_Pheno %>% filter(BillExposed != "NA") %>% 
  filter(Tarsometatarsus != "NA") %>% 
  filter(MiddleToe != "NA") %>% 
  filter(Tail != "NA") %>% 
  filter(Country != "USA")
str(Myna_Pheno_filter)

```


## Size correcting
By taking the natural log of the var and size

```{r, include=FALSE}
Myna_Pheno_filter2 <- Myna_Pheno_filter %>% do(data.frame(., 
      BillExposed.SC = residuals(lm(log(BillExposed)~ log(Tarsometatarsus), data=.)),
      BillNasal.SC = residuals(lm(log(BillNasal)~ log(Tarsometatarsus), data=.)),
      BillDepth.SC = residuals(lm(log(BillDepth)~ log(Tarsometatarsus), data=.)),
      BillWidth.SC = residuals(lm(log(BillWidth)~ log(Tarsometatarsus), data=.)),
      MiddleToe.SC = residuals(lm(log(MiddleToe)~ log(Tarsometatarsus), data=.)),
      Wing.SC = residuals(lm(log(Wing)~ log(Tarsometatarsus), data=.)),
      Tail.SC = residuals(lm(log(Tail)~ log(Tarsometatarsus), data=.))
      ))


str(Myna_Pheno_filter2)
```

## Scale
so all trait contribute equally

```{r, include=FALSE}
Myna_Pheno_filter2_scale <- Myna_Pheno_filter2 %>% mutate(
     Tarsus_scale = (Tarsometatarsus - mean(Tarsometatarsus)) / sd(Tarsometatarsus),
     BillExp_scale = (BillExposed.SC - mean(BillExposed.SC,na.rm=TRUE)) /sd(BillExposed.SC,na.rm=TRUE),
     BillNas_scale = (BillNasal.SC - mean(BillNasal.SC,na.rm=TRUE)) / sd(BillNasal.SC,na.rm=TRUE),
     BillDep_scale = (BillDepth.SC - mean(BillDepth.SC,na.rm=TRUE)) / sd(BillDepth.SC,na.rm=TRUE),
     BillWid_scale = (BillWidth.SC - mean(BillWidth.SC,na.rm=TRUE)) / sd(BillWidth.SC,na.rm=TRUE),
     Toe_scale = (MiddleToe.SC - mean(MiddleToe.SC,na.rm=TRUE)) / sd(MiddleToe.SC,na.rm=TRUE),
     Wing_scale = (Wing.SC - mean(Wing.SC,na.rm=TRUE)) / sd(Wing.SC,na.rm=TRUE),
     Tail_scale = (Tail.SC - mean(Tail.SC,na.rm=TRUE)) / sd(Tail.SC,na.rm=TRUE))

str(Myna_Pheno_filter2_scale)
```


#Dispersion Analysis

## Preparing for dispersion analysis

```{r, include=FALSE}
groupings <- as.factor(Myna_Pheno_filter2_scale$Country)

dispersion <- gowdis(Myna_Pheno_filter2_scale[,41:45])
str(dispersion)

betadisp <- betadisper(dispersion, groupings, type = c("centroid"), bias.adjust = FALSE, sqrt.dist = TRUE, add = FALSE)
dist.ALL <- betadisp$distances
str(dist.ALL)

FDis_MASTER.dist<-cbind(Myna_Pheno_filter2_scale[,c(9,17,18,12)],betadisp$distances)
colnames(FDis_MASTER.dist)[5] <- "distance"
str(FDis_MASTER.dist)
#not going to correct distance because it doesn't change results because the underlying vars are already log transformed

```


### LMER4 for dispersion
https://ourcodingclub.github.io/tutorials/mixed-models/
LOOK INTO: https://www.flutterbys.com.au/stats/tut/tut9.2a.html
https://stats.stackexchange.com/questions/27988/categorical-fixed-effect-w-3-levels-in-lmer

```{r}
full.lmer <- lmer(distance ~ Country + Sex + (1|Prov.State), data = FDis_MASTER.dist, REML = FALSE)
summary(full.lmer)
anova(full.lmer)
confint(full.lmer)
difflsmeans(full.lmer, test.effs = "Group")

pdf("Phenotypic_Dispersion_difflsmeans.pdf", width=18, height=8)
plot(difflsmeans(full.lmer, test.effs="Group"))
dev.off()


plot(full.lmer)
qqnorm(resid(full.lmer))
qqline(resid(full.lmer))
plot_model(full.lmer, type = "eff", terms = "State")
confint(full.lmer)

emmeans(full.lmer, pairwise ~ Sex | Country) # differences between sex for each State
emmeans(full.lmer, pairwise ~ Country | Sex) # differences between states for each Sex

data.nest.eff = as.data.frame(allEffects(full.lmer)[[1]])

dodge <- position_dodge(width=0.5)

ggplot(data.nest.eff, aes(y = fit, x = Country)) + 
  geom_pointrange(aes(ymin = lower,ymax = upper),position=dodge) + 
  scale_y_continuous("Phenotypic Dispersion (estimated marginal means)") + 
  theme_classic()+scale_color_manual(values=c("#E69F00", "#56B4E9")) +
  ylab("Population")+
  theme(legend.text = element_text(colour="black", size=12, 
                                   face="bold")) +
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))



```



### 2
### 2

## Load, filter, and check data

```{r, include = FALSE}
Myna_Pheno <- read.csv("myna_morphology_UOA.csv",stringsAsFactors=TRUE,sep=",")
str(Myna_Pheno)

#filter
Myna_Pheno_filter <- Myna_Pheno %>% filter(Bill.Exposed != "NA") %>% 
  filter(Tail != "NA")
str(Myna_Pheno_filter)

```


## Size correcting
By taking the natural log of the var and size

```{r, include=FALSE}
Myna_Pheno_filter2 <- Myna_Pheno_filter %>% do(data.frame(., 
      BillExposed.SC = residuals(lm(log(Bill.Exposed)~ log(Tarsus), data=.)),
      BillNasal.SC = residuals(lm(log(Bill.Nasal)~ log(Tarsus), data=.)),
      BillDepth.SC = residuals(lm(log(Bill.Depth)~ log(Tarsus), data=.)),
      BillWidth.SC = residuals(lm(log(Bill.Width)~ log(Tarsus), data=.)),
      MiddleToe.SC = residuals(lm(log(Middle.toe)~ log(Tarsus), data=.)),
      Wing.SC = residuals(lm(log(Wing)~ log(Tarsus), data=.)),
      Tail.SC = residuals(lm(log(Tail)~ log(Tarsus), data=.))
      ))


str(Myna_Pheno_filter2)
```

## Scale
so all trait contribute equally

```{r, include=FALSE}
Myna_Pheno_filter2_scale <- Myna_Pheno_filter2 %>% mutate(
     Tarsus_scale = (Tarsus - mean(Tarsus)) / sd(Tarsus),
     BillExp_scale = (BillExposed.SC - mean(BillExposed.SC,na.rm=TRUE)) /sd(BillExposed.SC,na.rm=TRUE),
     BillNas_scale = (BillNasal.SC - mean(BillNasal.SC,na.rm=TRUE)) / sd(BillNasal.SC,na.rm=TRUE),
     BillDep_scale = (BillDepth.SC - mean(BillDepth.SC,na.rm=TRUE)) / sd(BillDepth.SC,na.rm=TRUE),
     BillWid_scale = (BillWidth.SC - mean(BillWidth.SC,na.rm=TRUE)) / sd(BillWidth.SC,na.rm=TRUE),
     Toe_scale = (MiddleToe.SC - mean(MiddleToe.SC,na.rm=TRUE)) / sd(MiddleToe.SC,na.rm=TRUE),
     Wing_scale = (Wing.SC - mean(Wing.SC,na.rm=TRUE)) / sd(Wing.SC,na.rm=TRUE),
     Tail_scale = (Tail.SC - mean(Tail.SC,na.rm=TRUE)) / sd(Tail.SC,na.rm=TRUE))

str(Myna_Pheno_filter2_scale)
```


#Dispersion Analysis

## Preparing for dispersion analysis

```{r, include=FALSE}
groupings <- as.factor(Myna_Pheno_filter2_scale$Location)

dispersion <- gowdis(Myna_Pheno_filter2_scale[,19:26])
str(dispersion)

betadisp <- betadisper(dispersion, groupings, type = c("centroid"), bias.adjust = FALSE, sqrt.dist = TRUE, add = FALSE)
dist.ALL <- betadisp$distances
str(dist.ALL)

FDis_MASTER.dist<-cbind(Myna_Pheno_filter2_scale[,c(2,3,4)],betadisp$distances)
colnames(FDis_MASTER.dist)[4] <- "distance"
str(FDis_MASTER.dist)
#not going to correct distance because it doesn't change results because the underlying vars are already log transformed

```


### LMER4 for dispersion
https://ourcodingclub.github.io/tutorials/mixed-models/
LOOK INTO: https://www.flutterbys.com.au/stats/tut/tut9.2a.html
https://stats.stackexchange.com/questions/27988/categorical-fixed-effect-w-3-levels-in-lmer

```{r}
myna.anova<-aov(distance~Location,data=FDis_MASTER.dist)
summary(myna.anova)
anova(myna.anova)

plot(myna.anova)

TukeyHSD(myna.anova)


full.lmer <- lmer(distance ~ Country*Sex + (1|Prov.State), data = FDis_MASTER.dist, REML = FALSE)
summary(full.lmer)
anova(full.lmer)
confint(full.lmer)
difflsmeans(full.lmer, test.effs = "Group")

pdf("Phenotypic_Dispersion_difflsmeans.pdf", width=18, height=8)
plot(difflsmeans(full.lmer, test.effs="Group"))
dev.off()


plot(full.lmer)
qqnorm(resid(full.lmer))
qqline(resid(full.lmer))
plot_model(full.lmer, type = "eff", terms = "State")
confint(full.lmer)

emmeans(full.lmer, pairwise ~ Sex | Country) # differences between sex for each State
emmeans(full.lmer, pairwise ~ Country | Sex) # differences between states for each Sex

```




